---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(nycflights13)
theme_set(theme_minimal())
```


```{r}
today()
```

# Creating dates/times

## From strings
```{r}
ymd("2020-06-08")

mdy("06-08-2020")

dmy("08-06-2020")

# can also use unquoted numbers in these functions
ymd(20200608)

# date times just add _hms or similar to the end
ymd_hms("2020-06-08 13:11:59")
```


## From individual components

```{r}
flights%>%
  select(year, month, day, hour, minute)

#can create a date/time out of this using make_date() or make_datetime()
flights %>%
  select(year, month, day, hour, minute)%>%
  mutate(departure = make_datetime(year, month, day, hour, minute))

make_datetime_100 <- function(year, month, day, time){
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights_dt <- flights %>%
  filter(!is.na(dep_time), !is.na(arr_time))%>%
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  )%>%
  select(origin, dest, ends_with("delay"), ends_with("time"))
```

```{r}
flights_dt %>%
  ggplot(aes(dep_time))+
  geom_freqpoly(binwidth = 86400) # number of seconds in 1 day

flights_dt %>%
  filter(dep_time < ymd(20130102))%>%
  ggplot(aes(dep_time))+
  geom_freqpoly(binwidth = 600) # 600s = 10 minutes
```

## From other types

Switching between date and datetime uses as_date() and as_datetime()
```{r}
as_datetime(today())

as_date(now())
```

What happens if you parse a string that contains invalid dates?

```{r}
ymd(c("2020-01-01", "time"))
```
Warning is thrown that the item failed to parse and it parses as an NA

What does the tzone argument to today() do? Why is it important?

Allows you to specify with timezone you mean. Values of today can vary based on time zone so 
need to handle carefully.

Use the appropriate lubridate function to parse each of the following dates:

```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014

mdy(d1)
ymd(d2)
dmy(d3)
mdy(d4)
mdy(d5)
```

## Date-time Components
Can pull out individual components of date objects with the relevant year, month, mday, yday,
wday, hour, minute, second, etc function

```{r}
datetime <- ymd_hms("2016-07-08 12:34:56")
year(datetime)
month(datetime)
month(datetime, label = TRUE)

flights_dt %>%
  mutate(wday = wday(dep_time, label = TRUE))%>%
  ggplot(aes(x = wday))+
  geom_bar()
```

```{r}
dep <- flights_dt %>%
  mutate(minute = minute(dep_time))%>%
  group_by(minute)%>%
  summarise(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n())

dep%>%
  ggplot(aes(minute, avg_delay))+
  geom_line()

```

### Rounding
Use floor_date, round_date, ceiling_date to appropriately round a date to the nearby unit of time
```{r}
flights_dt %>%
  count(week = floor_date(dep_time, "week"))%>%
  ggplot(aes(week, n))+
  geom_line
```

### Setting components
Can use accessor functions to set components of a datetime
```{r}
datetime

year(datetime) <- 2020
datetime
```

```{r}
flights_dt %>%
  mutate(dep_hour = update(dep_time, yday = 1))%>%
  View()
```


How does the distribution of flight times within a day change over the course of the year?
```{r}
flights_dt%>%
  filter(!is.na(dep_time))%>%
  mutate(dep_hour = update(dep_time, yday = 1))%>%
  mutate(month = factor(month(dep_time)))%>%
  ggplot(aes(dep_hour, color = month))+
  geom_freqpoly(aes(y = ..density..),binwidth = 60*60)
```

Compare dep_time, sched_dep_time and dep_delay. Are they consistent? Explain your findings.
```{r}
flights_dt %>%
  mutate(dep_time_ = sched_dep_time + dep_delay * 60)%>%
  filter(dep_time != dep_time_)%>%
  select(dep_time_, dep_time, sched_dep_time, dep_delay)
```

Compare air_time with the duration between the departure and arrival. Explain your findings. (Hint: consider the location of the airport.)
```{r}
flights_dt%>%
  mutate(air_time_ = as.numeric(arr_time - dep_time),
         air_time_mins  = air_time,
         diff = air_time_ - air_time_mins)%>%
  select(origin, dest, air_time_, air_time_mins, diff)
```

How does the average delay time change over the course of a day? Should you use dep_time or sched_dep_time? Why?

```{r}
flights_dt %>%
  mutate(sched_dep_hour = hour(sched_dep_time))%>%
  group_by(sched_dep_hour)%>%
  summarise(dep_delay = mean(dep_delay))%>%
  ggplot(aes(x = sched_dep_hour, y = dep_delay))+
  geom_point()+
  geom_smooth()
```

On what day of the week should you leave if you want to minimise the chance of a delay?

```{r}
flights_dt %>%
  mutate(weekday = wday(sched_dep_time, abbr = TRUE, label = TRUE))%>%
  group_by(weekday)%>%
  summarise(dep_delay = mean(dep_delay))%>%
  mutate(weekday = fct_reorder(weekday, dep_delay))%>%
  ggplot(aes(x = weekday, y = dep_delay))+
  geom_col()+
  coord_flip()
```

What makes the distribution of diamonds$carat and flights$sched_dep_time similar?

```{r}
diamonds%>%
  ggplot(aes(carat %% 1 * 100))+
  geom_histogram(binwidth = 1)

flights_dt%>%
  ggplot(aes(minute(sched_dep_time)))+
  geom_histogram(binwidth = 1)
```

Confirm my hypothesis that the early departures of flights in minutes 20-30 and 50-60 are caused by scheduled flights that leave early. Hint: create a binary variable that tells you whether or not a flight was delayed.

```{r}
flights_dt %>%
  mutate(minute = minute(dep_time),
         early = dep_delay < 0)%>%
  group_by(minute)%>%
  summarise(
    early = mean(early, na.rm = TRUE),
    n = n()
  )%>%
  ggplot(aes(minute, early))+
  geom_line()
```


## Time spans

### Durations: represent an exact number of seconds

```{r}
a_age <- today() - ymd(19850712)
```
This is a difftime object. Lubridate provides a duration function that always uses seconds

```{r}
as.duration(a_age)
```

Bunch of duration constructors
```{r}
dseconds(15)
dminutes(10)
dhours(c(12, 24))
ddays(0:5)
dweeks(3)
dyears(1)
```

Can add and multiply durations
```{r}
2*dyears(1)
dyears(1) + dweeks(12) + dhours(15)

tomorrow <- today() + ddays(1)
last_year <- today() -dyears(1)
```

Because durations are exact number of seconds, you can sometimes get unexpected results--eg around daylight savings time.
```{r}
one_pm <- ymd_hms("2016-03-12 13:00:00", tz = "America/New_York")
one_pm
one_pm + ddays(1)

```


### Periods: represent human units like weeks and months
```{r}
one_pm
one_pm + days(1)
```

Similar constructors for periods, but they are human friendly terms

```{r}
seconds(15)
minutes(10)
hours(c(12, 24))
days(7)
months(1:6)
weeks(3)
years(1)

10*(months(6) + days(1))

days(50) + hours(25) + minutes(2)
```

Generally, compared to durations, periods are more likely to do what you expect.

```{r}
#leap year
ymd("2016-01-01") + dyears(1)

ymd("2016-01-01") + years(1)
```


Using periods to fix flights data: some planes appear to have arrived at dest before they departed NYC
```{r}
flights_dt %>%
  filter(arr_time < dep_time)
```

Really these are overnights flights.

```{r}
flights_dt <- flights_dt %>%
  mutate(
    overnight = arr_time < dep_time,
    arr_time = arr_time + days(overnight * 1),
    sched_arr_time = sched_arr_time + days(overnight * 1)
  )
```

### Intervals: represent starting and ending point

```{r}
dyears(1)/ddays(365)

years(1)/days(365)
```
Interval is a duration with a starting point--this precision allows you to determine exactly how long something was.

```{r}
next_year <- today() + years(1)

(today() %--% next_year) /ddays(1)

(today() %--% next_year) / days(1)
```

Choosing between duration, periods, and intervals? Choose the simplest data structure that solves your problem.

Why is there months() but no dmonths()?

Months are a human concept with no strictly defined value--some have 28 days, others 31. Since durations are exact
second values, dmonths wouldn't make sense.

Explain days(overnight * 1) to someone who has just started learning R. How does it work?
overnight is a boolean value of true or false. If the value of overnight is true, then the flight was an overnight flight, 
and the value of overnight is 1(true). So you would add one day to the value. If overnight is false(ie the flight wasn't 
an overnight flight), it's value is 0 and you add days(0*1) or zero days to the value.

Create a vector of dates giving the first day of every month in 2015. Create a vector of dates giving the first day of every month in the current year.

```{r}
ymd("2015-01-01") + months(0:11)

floor_date(today(), unit = 'year') + months(0:11)
```

Write a function that given your birthday (as a date), returns how old you are in years.

```{r}
my_age <- function(birthday){
  (birthday %--% today()) / years(1)
}
```

Why can’t (today() %--% (today() + years(1))) / months(1) work?

```{r}
(today() %--% (today() + years(1)))/ months(1)

# Do something like these instead
(today() %--% (today() + years(1))) %/% months(1)
(today() %--% (today() + years(1))) / days(30)


```
Dividing a duration by a period with inconsistent lengths. This won't always work like you expect and should be avoided.


## Time Zones
