---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


```{r}
x1 <- c("Dec", "Apr", "Jan", "Mar")
x2 <- c("Dec", "Apr", "Jam", "Mar")

```

To create a factor you must start by creating a list of the valid levels.

```{r}
month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun",
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

y1 <- factor(x1, levels = month_levels)

sort(y1)
```

Values not in the set are silently converted to NA

```{r}
y2 <- factor(x2, levels = month_levels)
```

Set the order of the levels to match the order of first appearance in the data
```{r}
f1 <- factor(x1, levels = unique(x1))
f1

# can do the same thing after the fact with fct_inorder()
f2 <- x1 %>%
  factor()%>%
  fct_inorder()
```

Access the set of valid levels with levels()
```{r}
levels(f2)
```

```{r}
gss_cat

gss_cat%>%
  count(race)

# force ggplot to display factor levels that don't have any values with scale_x_discrete(drop = FALSE)
gss_cat %>%
  ggplot(aes(race))+
  geom_bar()+
  scale_x_discrete(drop = FALSE)
```



Explore the distribution of rincome (reported income). What makes the default bar chart hard to understand? How could you improve the plot?

```{r}
gss_cat %>%
  ggplot(aes(rincome))+
  geom_bar()+
  coord_flip()
```
Default bar chart you can't see the axis values so you can flip those. Also, the ordering makes the graph tough to read.

What is the most common relig in this survey? What’s the most common partyid?
```{r}
# protestant
gss_cat %>%
  count(relig, sort = TRUE)

# independent
gss_cat %>%
  count(partyid, sort = TRUE)
```

Which relig does denom (denomination) apply to? How can you find out with a table? How can you find out with a visualisation?
```{r}
levels(gss_cat$denom)
```


## Modifying Factor Order

```{r}
relig_summary <- gss_cat %>%
  group_by(relig)%>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

relig_summary%>%
  ggplot(aes(tvhours, relig))+
  geom_point()

# better version
relig_summary%>%
  ggplot(aes(tvhours, fct_reorder(relig, tvhours)))+
  geom_point()

# better still
relig_summary%>%
  mutate(relig = fct_reorder(relig, tvhours))%>%
  ggplot(aes(tvhours, relig))+
  geom_point()
```


Shouldn't arbitrarily reorder variables that have and order though--like income. Try to only reorder
factors whose levels are arbitrary
```{r}
rincome_summary <- gss_cat%>%
  group_by(rincome)%>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

rincome_summary%>%
  mutate(rincome = fct_reorder(rincome, age))%>%
  ggplot(aes(age, rincome))+
  geom_point()
```

Can reorder a single level if it makes sense--for instance 'Not applicable' isn't like the other income variables
so it can makes sense to relevel it.
```{r}
rincome_summary%>%
  mutate(rincome = fct_relevel(rincome, "Not applicable"))%>%
  ggplot(aes(age, rincome))+
  geom_point()
```


fct_reorder2() reorders factor by the y values associated with the largest x values. 

```{r}
by_age <- gss_cat %>%
  filter(!is.na(age))%>%
  count(age, marital)%>%
  group_by(age)%>%
  mutate(prop = n/sum(n))

by_age %>%
  ggplot(aes(age, prop, color = marital))+
  geom_line(na.rm = TRUE)
```
This graph is tough to read because the lines don't match up with the legend color order

```{r}
by_age %>%
  ggplot(aes(age, prop, color = fct_reorder2(marital, age, prop)))+
  geom_line(na.rm = TRUE)+
  labs(color = 'Marital')
```

Can also use fct_infreq() to reorder by frequency easily
```{r}
gss_cat%>%
  mutate(marital = marital %>% fct_infreq() %>% fct_rev())%>%
  ggplot(aes(marital))+
  geom_bar()
```


There are some suspiciously high numbers in tvhours. Is the mean a good summary?
```{r}
gss_cat %>%
 filter(!is.na(tvhours))%>%
  ggplot(aes(tvhours))+
  geom_histogram(binwidth = 1)
```
Median may be more appropriate, but it depends on your goal.

For each factor in gss_cat identify whether the order of the levels is arbitrary or principled.
```{r}
gss_cat%>%
  select_if(is.factor)%>%
  map(., levels)
```

Why did moving “Not applicable” to the front of the levels move it to the bottom of the plot?


## Modifying factor levels

Modifying factor levels is the more powerful version of changing the orders of levels
fct_recode() is your first tool to do this

```{r}
gss_cat %>%
  count(partyid)

gss_cat %>%
  mutate(partyid = fct_recode(partyid,
    "Republican, strong"    = "Strong republican",
    "Republican, weak"      = "Not str republican",
    "Independent, near rep" = "Ind,near rep",
    "Independent, near dem" = "Ind,near dem",
    "Democrat, weak"        = "Not str democrat",
    "Democrat, strong"      = "Strong democrat"
  ))%>%
    count(partyid)
```

Can combine groups into one level. Use this with great care or you will end up combining groups that
are truly different.
```{r}
gss_cat %>%
  mutate(partyid = fct_recode(partyid,
    "Republican, strong"    = "Strong republican",
    "Republican, weak"      = "Not str republican",
    "Independent, near rep" = "Ind,near rep",
    "Independent, near dem" = "Ind,near dem",
    "Democrat, weak"        = "Not str democrat",
    "Democrat, strong"      = "Strong democrat",
    "Other"                 = "No answer",
    "Other"                 = "Don't know",
    "Other"                 = "Other party"
  )) %>%
  count(partyid)
```

Can use fct_collapse() to collapse multiple levels
```{r}
gss_cat %>%
  mutate(partyid = fct_collapse(partyid,
                                other = c("No answer", "Don't know", "Other party"),
                                rep = c("Strong republican", "Not str republican"),
                                ind = c("Ind,near rep", "Independent", "Ind,near dem"),
                                dem = c("Not str democrat", "Strong democrat")
  ))%>%
  count(partyid)
```

If you want to lump together small groups to make a plot or table simpler use fct_lump()

```{r}
gss_cat %>%
  mutate(relig = fct_lump(relig))%>%
  count(relig)

gss_cat %>%
  mutate(relig = fct_lump(relig, n = 10))%>%
  count(relig, sort = TRUE)

```



How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?
```{r}
gss_cat %>%
  mutate(partyid = fct_collapse(partyid,
                                other = c("No answer", "Don't know", "Other party"),
                                rep = c("Strong republican", "Not str republican"),
                                ind = c("Ind,near rep", "Independent", "Ind,near dem"),
                                dem = c("Not str democrat", "Strong democrat")
  ))%>%
  count(year, partyid)%>%
  group_by(year)%>%
  mutate(p = n/sum(n))%>%
  ggplot(aes(year, p, color = fct_reorder2(partyid, year, p)))+
  geom_line()+
  geom_point()
```


How could you collapse rincome into a small set of categories?
```{r}
gss_cat%>%
  mutate(rincome = fct_collapse(rincome,
                                Unknown = c("No answer", "Don't know", "Refused", "Not applicable"),
                                "Less than $5000" = c("Lt $1000", str_c(
                                  "$", c("1000", "3000", "4000"),
                                  " to ", c("2999", "3999", "4999")
                                )),
                                "$5000 to 10000" = str_c(
                                  "$", c("5000", "6000", "7000", "8000"),
                                  " to ", c("5999", "6999", "7999", "9999")
                                )))%>%
  ggplot(aes(x = rincome))+
  geom_bar()+
  coord_flip()

```

