---
title: "functions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## When to write a function?

Consider it once you've copied and pasted a block of code more than twice. Eg

```{r}
df <- tibble::tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
# going to function-ize this
df$a <- (df$a - min(df$a, na.rm = TRUE))/(max(df$a, na.rm = TRUE) - min(df$a, na.rm = TRUE))
```

First, look at what you want to make a function of--how many inputs does it have? One in the example above.
There is also duplication in the code--calculating a range which we can do in one step. Put those intermediate
calculations into well named variables. In total, the function could end up looking like this.
```{r}
rescale01 <- function(x){
  rng <- range(x, na.rm = TRUE)
  (x - rng[1])/(rng[2] - rng[1])
}

rescale01(df)
```

Three key steps to creating a new function:
1) Name: need to pick an evocative name for the function- above we chose rescale01 because the function rescales values
between 0 and 1.
2) Arguments: List the inputs(or arguments) inside function(x, y, z)
3) Body: code that actually is doing your work goings inside the function(){my code}

Exercises

1) Why is TRUE not a parameter to rescale01()? What would happen if x contained a single missing value, and na.rm was FALSE?
```{r}
x <- c(1:10, NA)

rescale01(x)
```

If you allow na.rm to be FALSE, a single NA value makes the entire input vector NA's.

2) In the second variant of rescale01(), infinite values are left unchanged. Rewrite rescale01() so that -Inf is mapped to 0, and Inf is mapped to 1.
```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

x <- c(1:10, Inf, -Inf)

rescale01(x)

rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  scaled <- (x - rng[1]) / (rng[2] - rng[1])
  scaled[scaled == Inf] <- 1
  scaled[scaled == -Inf] <- 0
  scaled
}
```

3) Practice turning the following code snippets into functions. Think about what each function does. What would you call it? How many arguments does it need? Can you rewrite it to be more expressive or less duplicative?

```{r}
# this is calculating the proportion of values in the vector that are NA
mean(is.na(x))

prop_na <- function(x){
  mean(is.na(x))
}

# this makes a vector sum to one
x / sum(x, na.rm = TRUE)

sum_to_one <- function(x, na.rm = FALSE){
  x / sum(x, na.rm = na.rm)
}

# this calculates coeficient of variation
sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)

coef_var <- function(x, na.rm = FALSE){
  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
}

coef_var(1:5)
coef_var(c(1:5, NA))
coef_var(c(1:5, NA), na.rm = TRUE)
```

4) Write your own functions to compute the variance and skewness of a numeric vector.

```{r}
variance <- function(x, na.rm = TRUE){
  n <- length(x)
  m <- mean(x, na.rm = TRUE)
  sq_error <- (x - m)^2
  sum(sq_error)/(n-1)
}

skewness <- function(x, na.rm = TRUE){
  n <- length(x)
  m <- mean(x, na.rm = TRUE)
  numerator <- sum((x - m)^3)/(n - 2)
  numerator/variance(x)^(3/2)
  
}
```

5) Write both_na(), a function that takes two vectors of the same length and returns the number of positions that have an NA in both vectors.

```{r}
x <- c(1:5, NA)
y <- c(6:10, NA)

both_na <- function(x, y){
  sum(is.na(x) & is.na(y))
}

both_na(
  c(NA, NA, 1, 2),
  c(NA, 1, NA, 2)
)

both_na(
  c(NA, NA, 1, 2, NA, NA, 1),
  c(NA, 1, NA, 2, NA, NA, 1)
)
```

6) What do the following functions do? Why are they useful even though they are so short?

Useful because they are much clearer in what they are doing than the underlying code
```{r}
# this takes a file/document path and tells whether it is a directory or not.
is_directory <- function(x) file.info(x)$isdir

# tests whether a file is readable--has the readable permissions. Value of 0 means it's readable. -1 means it's not
is_readable <- function(x) file.access(x, 4) == 0
```

7) Read the complete lyrics to “Little Bunny Foo Foo”. There’s a lot of duplication in this song. Extend the initial piping example to recreate the complete song, and use functions to reduce the duplication.

```{r}

```


Exercises

1) Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
has_prefix <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
drop_last <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
expand <- function(x, y) {
  rep(y, length.out = length(x))
}
```



1) What’s the difference between if and ifelse()? Carefully read the help and construct three examples that illustrate the key differences.
```{r}
ifelse(1:5 > 2, TRUE, FALSE)

if(1:5 > 2){
  TRUE
}else{
  FALSE
}
```

2) Write a greeting function that says “good morning”, “good afternoon”, or “good evening”, depending on the time of day. (Hint: use a time argument that defaults to lubridate::now(). That will make it easier to test your function.)

```{r}
greetings <- function(time = lubridate::now()){
  if(lubridate::hour(time) < 12){
    print("Good Morning")
  }else if(lubridate::hour(time) >= 12 & lubridate::hour(time) < 19){
    print("Good Afternoon")
  }else{
    print("Good Evening")
  }
}
```

3) Implement a fizzbuzz function. It takes a single number as input. If the number is divisible by three, it returns “fizz”. If it’s divisible by five it returns “buzz”. If it’s divisible by three and five, it returns “fizzbuzz”. Otherwise, it returns the number. Make sure you first write working code before you create the function.

```{r}
fizzbuzz <- function(number){
  # check if number if a valid input
  stopifnot(length(number) == 1)
  stopifnot(is.numeric(number))
  
  if(number %% 3 == 0 && number %% 5 == 0){
    return("Fizzbuzz")
  }else if (number %% 5 == 0){
    return("Buzz")
  }else if(number %% 3 == 0){
    return("Fizz")
  } else{
    return(as.character(number))
  }
  
}
library(dplyr)
fizzbuzz_vec <- function(x){
  case_when(!(x %% 5) & !(x %% 3) ~ "fizzbuzz",
            !(x %% 5) ~ "Buzz",
            !(x %% 3) ~ "Fizz",
            TRUE ~ as.character(x))
}

fizzbuzz_vec2 <- function(x){
  y <- as.character(x)
  
  y[!(x %% 3)] <- "fizz"
  y[!(x %% 5)] <- "buzz"
  y[!(x %%3) & !(x %% 5)] <- "fizzbuzz"
  
  y
}
```

4) How could you use cut() to simplify this set of nested if-else statements?

```{r}
if (temp <= 0) {
  "freezing"
} else if (temp <= 10) {
  "cold"
} else if (temp <= 20) {
  "cool"
} else if (temp <= 30) {
  "warm"
} else {
  "hot"
}

cut(-20:40, breaks = c(-Inf, 0, 10, 20, 30, Inf), labels = c("freezing", "cold", "cool", "warm", "hot"))
```

5) What happens if you use switch() with numeric values?
```{r}
?switch

# if numeric, it will return the nth argument from ...
switch(2, "blue", "red", "green")
```

6) What does this switch() call do? What happens if x is “e”?

```{r}
x <- "e"
switch(x, 
  a = ,
  b = "ab",
  c = ,
  d = "cd"
)
```

Exercises
1) What does commas(letters, collapse = "-") do? Why?
```{r}
commas <- function(...) stringr::str_c(..., collapse = ", ")

```
collapse is passed to commas through .... It already has a collapse argument so you are supplying two.
str_c(letters, collapse = "-", collapse = ", "). This doesn't work. 


2)It’d be nice if you could supply multiple characters to the pad argument, e.g. rule("Title", pad = "-+"). Why doesn’t this currently work? How could you fix it?

```{r}
rule <- function(..., pad = "-") {
  title <- paste0(...)
  width <- getOption("width") - nchar(title) - 5
  cat(title, " ", stringr::str_dup(pad, width), "\n", sep = "")
}

library(tidyverse)
rule2 <- function(..., pad = "-"){
  title <- paste0(...)
  width <- getOption("width") - nchar(title) - 5
  padding <- str_dup(
    pad,
    ceiling(width / str_length(title))
  )%>%
    str_trunc(width)
  
  cat(title, " ", padding, "\n", sep = "")
}
```
Current rule function assumes a pad character will be a single character. So the output is double length. Can fix
this duplication with str_trunc()


3) What does the trim argument to mean() do? When might you use it?

```{r}
?mean
```

trim symmetrically trims the vector with a fraction of observations deleted from each end of the vector before computing the mean.
This is useful for calculating a measure of central tendency that's robust to outliers.


4) The default value for the method argument to cor() is c("pearson", "kendall", "spearman"). What does that mean? What value is used by default?
```{r}
?cor
```
Means the method can take one of those 3 values. The default being the first value.

