---
title: "Iteration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```



## For Loops

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

output <- vector("double", ncol(df))
for(i in seq_along(df)){
  output[[i]] <- median(df[[i]])
}

```

Every for loop has:
1) Output: Before you start the loop, you need to allocate space for the output. If you don't do this and instead
grow the for loop at each iteration using c(), the loop will be very slow.

2) Sequence: i in seq_length(df). This determines what you are looping over.

3) Body: output[[i]] <- median(df[[i]]). Code that does the work you want done. Run repeatedly with a different value for i 
each time.

Exercises

1)
Compute the mean of every column in mtcars.
```{r}
means <- vector("double", ncol(mtcars))
for(i in seq_along(mtcars)){
  means[[i]] <- mean(mtcars[[i]])
}
means
```
Determine the type of each column in nycflights13::flights.
```{r}
col_type <- vector("list", ncol(nycflights13::flights))
names(col_type) <- names(nycflights13::flights)
for(i in seq_along(nycflights13::flights)){
  col_type[[i]] <- class(nycflights13::flights[[i]])
}
col_type
glimpse(nycflights13::flights)
```
Compute the number of unique values in each column of iris.
```{r}
output <- vector("double", ncol(iris))
for(i in seq_along(iris)){
  output[[i]] <- length(unique(iris[[i]]))
}
output
```
Generate 10 random normals from distributions with means of -10, 0, 10, and 100.

```{r}
means <- c(-10, 0, 10, 100)
output <- vector("list", length(means))
for(i in seq_along(means)){
  output[[i]] <- rnorm(10, mean = means[[i]])
}

output

# rnorm recycles mean argument, so this isn't needed
matrix(rnorm(10 * length(means), mean = means), ncol = 10)
```

2) Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors:

```{r}
# out <- ""
# for (x in letters) {
#   out <- stringr::str_c(out, x)
# }
str_c(letters, collapse = "")

x <- sample(100)
sd <- 0
for (i in seq_along(x)) {
  sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))

sd(x)


x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}

cumsum(x)
```

3) Combine your function writing and for loop skills:

Write a for loop that prints() the lyrics to the children’s song “Alice the camel”.

Convert the nursery rhyme “ten in the bed” to a function. Generalise it to any number of people in any sleeping structure.

Convert the song “99 bottles of beer on the wall” to a function. Generalise to any number of any vessel containing any liquid on any surface.
```{r}

```

4) It’s common to see for loops that don’t preallocate the output and instead increase the length of a vector at each step:


```{r}
add_to_vector <- function(n){
  output <- vector("integer", 0)
  
  for(i in seq_len(n)){
    output <- c(output, i)
  }
  output
}

add_to_vector2 <- function(n){
  output <- vector("integer", n)
  for(i in seq_len(n)){
    output[[i]] <- i
  }
  output
}

timings <- microbenchmark(add_to_vector(10000), add_to_vector2(10000), times = 10)
```



How does this affect performance? Design and execute an experiment.


## For Loop Variations

### Modifying an existing object
```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

rescale01 <- function(x){
  rng <- range(x, na.rm = TRUE)
  (x - rng[1])/ (rng[2] - rng[1])
}

for(i in seq_along(df)){
  df[[i]] <- rescale01(df[[]])
}
```

### Looping over names or values, instead of indices
Looping over numeric indices is the most general way we loop. Can also
1) Loop over the elements: for (x in xs)
2) Loop over the names: for (nm in names(xs)). Can use this to access the name for a plot title or file name.

Iteration over numeric indices is still most general because using this you can access name and value
```{r, eval = FALSE}
for (i in seq_along(x)){
  name <- names(x)[[i]]
  value <- x[[i]]
}
```


### Handling outputs of unknown length

You may not always know how long the output will be. Initial thought might be to grown the vector progressively,
but this is O(n^2), so it's very inefficient

Better solution is to save the results in a list and then combine into a single vector after the loop is done:
```{r}
means <- c(0, 1, 2)

out <- vector('list', length(means))
for(i in seq_along(means)){
  n <- sample(100, 1)
  out[[i]] <- rnorm(n, means[[i]])
}

str(out)
str(unlist(out))

# stricter option for flattening a list is purrr::flatten_dbl() which will throw an error if the input isn't a list of doubles
```

This sort of pattern occurs in many places:
1) May be generating a long string. Instead of pasting at each iteration, save the output to a character vector and collapse it later.

2) May generate a big data frame. Instead of sequential rbind() at each iteration, save the output in a list and then bind_rows() to combine
to a single df.

When you see this pattern, switch to a more complex result object and then combine in one step at the end

### Handling sequences of unknown length

When you don't know how long the input sequence may run, that's a situation for a while loop.
```{r}
while (condition){
  #body
}
```
Any for loop can be a while loop. Not every while loop can be a for loop.
```{r}
for(i in seq_along(x)){
  # body
}

# equivalent to
i <- 1
while(i <= length(x)){
  # body
  i <- i + 1
}
```

While loop to figure out how many tries it takes to get 3 heads in a row
```{r}
flip <- function() sample(c("T", "H"), 1)

flips <- 0
nheads <- 0

while (nheads < 3){
  if(flip() == "H"){
    nheads <- nheads + 1
  } else{
    nheads <- 0
  }
  flips <- flips + 1
}

flips
```

Exercises
1) Imagine you have a directory full of CSV files that you want to read in. You have their paths in a vector, files <- dir("data/", pattern = "\\.csv$", full.names = TRUE), and now want to read each one with read_csv(). Write the for loop that will load them into a single data frame.

```{r}
df_list <- vector('list', length(files))

for(i in seq_along(files)){
  df_list[[i]] <- read_csv(files[[i]])
}

df <- bind_rows(df_list)
```

2) What happens if you use for (nm in names(x)) and x has no names? What if only some of the elements are named? What if the names are not unique?
```{r}
x <- c(1, 2, 3)
print(names(x))
for(nm in names(x)){
  print(nm)
  print(x[[nm]])
}
```
It doesn't run the loop at all if there are no names.

```{r}
x <- c(a = 1, b = 2, 3)
print(names(x))
for(nm in names(x)){
  print(nm)
  print(x[[nm]])
}

```

With incomplete names, an error is thrown when you try to access an element without a name.

```{r}
x <- c(a = 1, a = 2, b = 3)
print(names(x))
for(nm in names(x)){
  print(nm)
  print(x[[nm]])
}
```
With duplicate names, the first element with that name is returned


3) Write a function that prints the mean of each numeric column in a data frame, along with its name. For example, show_mean(iris) would print:

```{r}
show_mean(iris)
#> Sepal.Length: 5.84
#> Sepal.Width:  3.06
#> Petal.Length: 3.76
#> Petal.Width:  1.20

show_mean <- function(x){
  for(nm in names(x)){
    if(is.numeric(x[[nm]])){
      cat(
        str_c(nm, ": ", format(mean(x[[nm]]), digits = 2, nsmall = 2)),
        "\n"
      )
    }
  }
}
```

4) What does this code do? How does it work?

```{r}
trans <- list( 
  disp = function(x) x * 0.0163871,
  am = function(x) {
    factor(x, labels = c("auto", "manual"))
  }
)
for (var in names(trans)) {
  mtcars[[var]] <- trans[[var]](mtcars[[var]])
}
```

This mutates the displacement and am column. It multiplies disp by .016 and am is made into a factor variable.


## For loops vs. functionals

Since R is a functional language, you can wrap for loops in a function and call that function instead of using the loop
directly.

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)


```

Exercises
2) Adapt col_summary() so that it only applies to numeric columns You might want to start with an is_numeric() function that returns a logical vector that has a TRUE corresponding to each numeric column.

```{r}
col_summary <- function(df, fun) {
  numeric_cols <- vector('logical', length(df))
  
  for(i in seq_along(df)){
    numeric_cols[[i]] <- is.numeric(df[[i]])
  }
  
  idxs <- which(numeric_cols)
  n <- sum(numeric_cols)
  
  out <- vector("double", n)
  for (i in seq_along(idxs)) {
    out[[i]] <- fun(df[[idxs[[i]]]])
  }
  names(out) <- names(df)[idxs]
  out
}
```

## Map functions

purrr family of functions are consistent. They:
1) take a vector as input
2) apply a function to each piece
3) return a new vector that's the same length(and has the same names) as the input

The type of the vector is determined by the suffix tot he map function

Chief benefit of using map functions is clarity, not really speed. If you write proper for loops they are plenty fast,
but map functions are typically easier to write and read.


```{r}
map_dbl(df, mean)
map_dbl(df, median)

df %>%
  map_dbl(mean)
```

* Purrr functions *are* implemented in C so they are a little faster than something like the col_summary function above.
* The function argument in purrr functions, .f, can be a formula, a character vector, or an integer vector.
* map_* uses ... to pass along additional arguments to .f each time it's called.
* map functions also preserve names.

### Shortcuts

Say you are fitting a linear model to each group in a dataset
```{r}
models <- mtcars%>%
  split(.$cyl)%>%
  map(function(df) lm(mpg ~ wt, data = df))
```

This is more verbose than you need to be with purrr. You can use a one sided formula in purrr. You also use . as a pronoun:
it refers to the current list element(same way that i refers to the current index in the for loop).
```{r}
models <- mtcars %>%
  split(.$cyl)%>%
  map(~lm(mpg ~ wt, data = .))
```

You will often want to extract summary statistics. You could do either of the options below.

```{r}
models %>%
  map(summary)%>%
  map_dbl(~.$r.squared)

models %>%
  map(summary)%>%
  map_dbl("r.squared")
```

Can also use an integer to select elements by position
```{r}
x <- list(list(1,2,3), list(4, 5, 6), list(7, 8, 9))
x %>%
  map_dbl(2)
```


Exercises

1) Write code that uses one of the map functions to:

Compute the mean of every column in mtcars.
```{r}
mtcars %>%
  map_dbl(mean)
```

Determine the type of each column in nycflights13::flights.
```{r}
nycflights13::flights%>%
  map_chr(typeof)
```

Compute the number of unique values in each column of iris.

```{r}
iris %>%
  map_int(~length(unique(.)))
```

Generate 10 random normals from distributions with means of -10, 0, 10, and 100.
```{r}
c(-10, 0, 10, 100)%>%
  map(~rnorm(n = 10, mean = .))
```

2) How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?

```{r}
iris%>%
  map_lgl(is.factor)
```

3) What happens when you use the map functions on vectors that aren’t lists? What does map(1:5, runif) do? Why?
```{r}
map(1:5, runif)
```
Can use map on any type of vector, not just lists. This loops through 1:5 and returns a list of random variables 
from the uniform distribution--first 1, then 2, then 3, ...

4) What does map(-2:2, rnorm, n = 5) do? Why? What does map_dbl(-2:2, rnorm, n = 5) do? Why?
```{r}
map(-2:2, rnorm, n = 5)

map_dbl(-2:2, rnorm, n = 5)
```

Returns a list, with each element containing 5 random  variables centered at -2:2.
The second expression throws an error because map_dbl requires the function it applies to each element to return 
a vector of length one.

To return a numeric vector you can use flatten_dbl
```{r}
map(-2:2, rnorm, n = 5)%>%
  flatten_dbl()
```

5) Rewrite map(x, function(df) lm(mpg ~ wt, data = df)) to eliminate the anonymous function.

```{r}
x %>%
  map(~lm(mpg ~ wt, data = .))
```


## Dealing with failure

When using map to repeat many operations, the chances are higher that one of the operations might fail. If this happens, you
get an error message and no output. To avoid this, you can use the function safely()--it takes a function and returns a modified version.

Ex.
```{r}
safe_log <- safely(log)
str(safe_log(10))

str(safe_log("a"))

# works with map
x <- list(1, 10, "a")

y <- x %>%
  map(safely(log))

str(y)
```

Easier to work with this sort of output if you have two lists here--one with errors, one with the output
```{r}
y <- y %>%
  transpose()

str(y)
```

Two additional adverbs in the safely vein: 

possibly(): similar to safely it always succeeds, but you give it a default value to return when there is an error.
```{r}
x <- list(1, 10, "a")

x %>%
  map_dbl(possibly(log, NA_real_))
```

quietly(): similar to safely, but instead of capturing the errors, it captures printed output, messages and warnings
```{r}
x <- list(1, -1)

x %>%
  map(quietly(log))%>%
  str()
```

## Mapping over multiple arguments

Mapping over multiple inputs uses map2() and pmap()

```{r}
mu <- list(5, 10, -3)

mu %>%
  map(rnorm, n = 5)%>%
  str()
```

Say you want to vary SD too. One way is iterating over indices, but this isn't super clear.
```{r}
sigma <- list(1, 5, 10)

seq_along(mu)%>%
  map(~rnorm(5, mu[[.]], sigma[[.]]))%>%
  str()
```

Instead, you can use map2
```{r}
map2(mu, sigma, rnorm, n = 5)%>%
  str()

```

Arguments that vary for each call come *before* the function. Arguments that are the *same* come after.

pmap() is an extention of this to p arguments--say you want to vary mean, sd, and number of samples. Pmap will use positional
matching if names are not supplied. This is confusing, so use named arguments. Arguments must be a list of vectors--that means you can,
and should as the code gets complicated--store the args in a data.frame or tibble
```{r}
n <- list(1,2,3)

args1 <- list(n = n, mean = mu, sd = sigma)
args1 %>%
  pmap(rnorm)

params <- tribble(
  ~mean, ~sd, ~n,
  5, 1, 1, 
  10, 5, 3,
  -3, 10, 5
)

params %>%
  pmap(rnorm)
```

### Invoking different functions

In addition to changing parameters, you can vary the function itself. This is handled with invoke_map()

```{r}
f <- c("runif", "rnorm", "rpois")

param <- list(
  list(min = -1, max = 1),
  list(sd = 5),
  list(lambda = 10)
)

invoke_map(f, param, n = 5)

# can use tribble to make matching up these pairs easier
sim <- tribble(
  ~f, ~params,
  "runif", list(min = -1, max = 1),
  "rnorm", list(sd = 5),
  "rpois", list(lambda = 10)
)

sim %>%
  mutate(sim = invoke_map(f, params, n = 10))%>%
  .$sim
```


## Walk
Walk is a map alternative when you are calling a function for its side effects, rather than its return value.

```{r}
x <- list(1, "a", 3)

x %>%
  walk(print)
```

Like map, walk2 and pwalk exist.

```{r}
plots <- mtcars %>%
  split(.$cyl)%>%
  map(~ggplot(., aes(mpg, wt)) + geom_point())

paths <- stringr::str_c(names(plots), ".pdf")
pwalk(list(paths, plots), ggsave, path = tempdir())
```

## Other types of for loops
There are a lot of other purrr functions that abstract other types of for loops. Some examples here so I can hopefully
remember a function exists and then look up how to use it.

### Predicate Functions

keep() and discard() keep elements of the input where the predicate is TRUE or FALSE respectively
```{r}
iris %>%
  keep(is.factor)%>%
  str()

iris %>%
  discard(is.factor)%>%
  str()
```

some() and every() determine if the predicate is true for any or for all of the elements

```{r}
x <- list(1:5, letters, list(10))

x %>%
  some(is_character)

x %>%
  every(is_character)

x %>% 
  every(is_vector)
```

detect() finds the 1st element where the predicate is true; detect_index() returns the position of the first true element

```{r}
x <- sample(10)
x
x %>%
  detect(~ . > 5)

x %>%
  detect_index(~ . > 5)
```


head_while() and tail_while() take elements from the start or end of a vector while a predicate is true

```{r}
x %>%
  head_while(~ . > 5)

x %>%
  tail_while(~ . > 5)
```

### Reduce and Accumulate

Sometimes you have a complex list you want to reduce to a simple list by repeatedly applying a function that reduces a pair to 
a singleton. EG: You want to apply a two-table dplyr verb to multiple tables. You might have a list of data frames, and you wanted
to reduce to a single data frame by joining elements together.

```{r}
dfs <- list(
  age = tibble(name = "John", age = 30),
  sex = tibble(name = c("John", "Mary"), sex = c("M", "F")),
  trt = tibble(name = "Mary", treatment = "A")
)

dfs %>%
  reduce(full_join)
```

Could be a list of vectors that you wish to find the intersection of

```{r}
vs <- list(
  c(1, 3, 5, 6, 10),
  c(1, 2, 3, 7, 8, 10),
  c(1, 2, 3, 4, 8, 9, 10)
)

vs %>%
  reduce(intersect)
```

Reduce takes a binary function(one with 2 primary inputs) and applies it repeatedly to a list until there is a single element left.

Accumulate() keeps all interim results

```{r}
x <- sample(10)
x

x %>%
  accumulate(`+`)
```


Exercises

1) Implement your own version of every() using a for loop. Compare it with purrr::every(). 
What does purrr’s version do that your version doesn’t?

```{r}
x <- list(1:5, letters, list(10))

my_every <- function(.x, .p, ...){
  for(el in .x){
    if(!.p(i, ...)){
      return(FALSE)
    }
  }
  TRUE
}

letters %>%
  my_every(is_character)
```

2) Create an enhanced col_summary() that applies a summary function to every numeric column in a data frame.

```{r}
col_summary <- function(.x, .f, ...){
  if(!some(.x, is.numeric)){
    return("No Numeric Columns")
  }
  
  num_cols <- keep(.x, is.numeric)
  
  num_cols %>%
    map(.f)
}
```

3) A possible base R equivalent of col_summary() is:

```{r}
col_sum3 <- function(df, f) {
  is_num <- sapply(df, is.numeric)
  df_num <- df[, is_num]

  sapply(df_num, f)
}
```


But it has a number of bugs as illustrated with the following inputs:

```{r}
df <- tibble(
  x = 1:3, 
  y = 3:1,
  z = c("a", "b", "c")
)
# OK
col_sum3(df, mean)
# Has problems: don't always return numeric vector
col_sum3(df[1:2], mean)
col_sum3(df[1], mean)
col_sum3(df[0], mean)
```

What causes the bugs?

sapply() doesn't guarantee a consistent return type.

