---
title: "Programming-pipes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(magrittr)
```


```{r}
diamonds <- ggplot2::diamonds

diamonds2 <- diamonds %>%
  dplyr::mutate(price_per_carat = price/carat)

pryr::object_size(diamonds)

pryr::object_size(diamonds2)

pryr::object_size(diamonds, diamonds2)
```


Using the pipe you can assemble something like 
```{r, eval = FALSE}
foo_foo %>%
  hop(through = forest)%>%
  scoop(up = field_mice)%>%
  bop(on = head)
```

Behind the scenes the pipe is assembling something like this though
```{r, eval = FALSE}
my_pipe <- function(.){
  . <- hop(., through = forest)
  . <- scoop(., up = field_mic)
  bop(., on = head)
}

my_pipe(foo_foo)
```

So pipes won't work for 2 classes of functions

1) Functions that use the current environment. assign() creates a new variable with given name in the 
current environment. Other functions that face this problem include get() and load().
```{r}
assign("x", 10)

# this doesn't work because the it assigns it to a temp environ used by %>%
# have to be explicit about the environment
"x" %>% assign(10)

# this would work
env <- environment()
"x" %>% assign(100, envir = env)
```

2) Functions that use lazy evaluation. In R, function arguments are only computed when the function uses them, not prior
to calling the function. The pipe computes each element in turn, so you can't rely on this behavior.
```{r}
tryCatch(stop("!"), error = function(e) "An error")

stop("!")%>%
  tryCatch(error = function(e) "An error")
```

A pretty big number of functions use this behavior in base R -- tryCatch, try, suppressMessages, suppressWarnigns, etc


### When NOT to use the pipe

Pipes are mainly useful for rewriting a fairly short linear sequence of operations. You should reach for another tool when:

* Your pipes are longer than, say, ten steps. In this case, intermediate objects with good names will make debugging easier
and make your code easier to understand because variable names communicate intent.

* You have multiple inputs or outputs. If there isn't one primary object being transformed, but two or more objects being combined
together, don't use the pipe.

* You are starting to think about a directed graph with a complex dependency structure. Pipes are fundamentally linear
so expressing complex relationships is typically confusing.


### Other magrittr tools
```{r}
rnorm(100)%>%
  matrix(ncol = 2)%>%
  plot()%>%
  str()

rnorm(100)%>%
  matrix(ncol = 2)%T>%
  plot()%>%
  str()

mtcars %>%
  cor(disp, mpg)

mtcars %$%
  cor(disp, mpg)

```


