---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(nycflights13)
```

```{r}
airlines
airports
planes
weather
```

```{r}
flights_latlon <- flights %>%
  inner_join(select(airports, origin = faa, origin_lat = lat, origin_lon = lon), 
             by = "origin")%>%
  inner_join(select(airports, dest = faa, dest_lat = lat, dest_lon = lon),
             by = "dest")

flights_latlon %>%
  slice(1:100)%>%
  ggplot(aes(
    x = origin_lon, xend = dest_lon,
    y = origin_lat, yend = dest_lat
  ))+
  borders("state")+
  geom_segment(arrow = arrow(length = unit(0.1, "cm")))+
  coord_quickmap()
```

```{r}
planes%>%
  count(tailnum)

weather %>%
  count(year, month, day, hour, origin)%>%
  filter(n > 1)

flights %>%
  count(year, month, day, tailnum)%>%
  filter(n > 1)
```

```{r}
flights %>%
  mutate(flight_id = row_number())%>%
  glimpse()

Lahman::Batting%>%
  glimpse()

Lahman::Batting%>%
  count(playerID, yearID, stint)%>%
  filter(n>1)

babynames::babynames%>%
  count(year, sex, name)%>%
  filter(n> 1)

nasaweather::atmos %>%
  count(lat,long, year, month)%>%
  filter(n > 1)

fueleconomy::vehicles%>%
  count(id)%>%
  filter(n > 1)

ggplot2::diamonds%>%
  count(x, y, z, price, clarity, cut)
```


## Mutating Joins
```{r}
flights2 <- flights %>%
  select(year:day, hour, origin, dest, tailnum, carrier)

flights2 %>%
  left_join(airlines, by = "carrier")

# base equivalent--or using base subsetting at least
flights2 %>%
  mutate(name = airlines$name[match(carrier, airlines$carrier)])
```


```{r}
x <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2", 
  3, "x3"
)

y <- tribble(
  ~key, ~val_y,
  1, "y1", 
  2, "y2", 
  4, "y3"
)

x %>%
  inner_join(y, by = 'key')
```


## Duplicate Keys

```{r}
x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     2, "x3",
     1, "x4"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2"
)

x %>%
  left_join(y, by = 'key')


x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     2, "x3",
     3, "x4"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2",
     2, "y3",
     3, "y4"
)

x %>%
  left_join(y, by = 'key')
```

## Defining key columns

```{r}
flights2 %>%
  left_join(weather)

flights2 %>%
  left_join(planes, by = 'tailnum')

flights2 %>%
  left_join(airports, c("dest" = "faa"))
```

Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays.
```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()

delays <- flights %>%
  group_by(dest)%>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE))%>%
  inner_join(airports, by = c("dest" = "faa"))


delays %>%
  ggplot(aes(lon, lat, color = avg_delay))+
  borders("state")+
  geom_point()+
  coord_quickmap()
```

Add the location of the origin and destination (i.e. the lat and lon) to flights

```{r}
airport_locs <- airports%>%
  select(faa, lat, lon)

flights %>%
  select(year:day, hour, origin, dest)%>%
  left_join(airport_locs, by = c("origin" = "faa"))%>%
  left_join(airport_locs, by = c("dest" = "faa"), suffix = c("_origin", "_dest"))
```

Is there a relationship between the age of a plane and its delays?

```{r}
flights_with_planes <- flights %>%
  left_join(planes, by = "tailnum", suffix = c("_flight", "_built"))

flights_with_planes %>%
  select(year_flight:sched_arr_time, year_built)%>%
  filter(!is.na(year_built))%>%
  mutate(plane_age = year_flight - year_built)%>%
  group_by(plane_age)%>%
  summarise(avg_delay = mean(dep_delay, na.rm = TRUE))%>%
  ggplot(aes(x = plane_age, y = avg_delay))+
  geom_point()+
  geom_smooth()

```

What weather conditions make it more likely to see a delay?

```{r}
flight_weather <- flights %>%
  inner_join(weather, by = c("origin" = "origin", "year" = "year", "month" = "month",
                             "day" = "day", "hour" = "hour"))

flight_weather%>%
  select(year:day, dep_delay, arr_delay, flight, tailnum, origin, temp, temp:visib)%>%
  group_by(visib)%>%
  summarise(avg_delay = mean(dep_delay, na.rm = TRUE))%>%
  ggplot(aes(visib, avg_delay))+
  geom_line()
```

What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather

```{r}
flights %>%
  filter(year == 2013, month == 6, day == 13)%>%
  group_by(dest)%>%
  summarise(delay = mean(arr_delay, na.rm = TRUE))%>%
  inner_join(airports, by = c('dest' = 'faa'))%>%
  ggplot(aes(y = lat, x = lon, size = delay, color = delay))+
  borders("state")+
  geom_point()+
  coord_quickmap()+
  scale_color_viridis_c()
```

What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)

```{r}
flights %>%
  filter(is.na(tailnum), !is.na(arr_time))

flights %>%
  anti_join(planes, by = "tailnum")%>%
  count(carrier, sort = TRUE)
```

Filter flights to only show flights with planes that have flown at least 100 flights.

```{r}
over_100 <- flights %>%
  filter(!is.na(tailnum))%>%
  count(tailnum, sort = TRUE)%>%
  filter(n >=100)

flights %>%
  semi_join(over_100, by = "tailnum")
```

Combine fueleconomy::vehicles and fueleconomy::common to find only the records for the most common models.

```{r}
fueleconomy::vehicles%>%
  semi_join(fueleconomy::common, by = c("make", "model"))
```

Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the weather data. Can you see any patterns?

```{r}
worst_hours <- flights %>%
  mutate(hour = sched_dep_time %/% 100)%>%
  group_by(origin, year, month, day, hour)%>%
  summarise(dep_delay = mean(dep_delay, na.rm = TRUE))%>%
  ungroup()%>%
  arrange(desc(dep_delay))%>%
  slice(1:48)

weather %>%
  semi_join(worst_hours, by = c(
    "origin", "year", "month", "day", "hour"
  ))

```


What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?

```{r}
anti_join(flights, airports, by = c("dest" = "faa"))
anti_join(airports, flights, by = c("faa" = "dest"))
```

First is flights that landed at an airport not in the FAA airport list of destinations. Second is airports that didn't have a single 
flight land at them.

You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above.

```{r}
planes_carriers <- flights%>%
  filter(!is.na(tailnum))%>%
  distinct(tailnum, carrier)

planes_carriers%>%
  count(tailnum)%>%
  filter(n > 1)
```


## Set operations
```{r}
df1 <- tribble(
  ~x, ~y,
   1,  1,
   2,  1
)
df2 <- tribble(
  ~x, ~y,
   1,  1,
   1,  2
)

intersect(df1, df2)
union(df1, df2)

setdiff(df1, df2)
setdiff(df2, df1)

```

