---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

tibble(
  x = 1:5,
  y = 1, 
  z = x ^ 2 + y
)

tibble(
  `:)` = "smile",
  ` ` = "space",
  `2000` = "number"
)


tribble(
  ~x, ~y, ~z,
  #--/--/----
  "a", 2, 3.6,
  "b", 1, 8.5
)
```


```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
#> # A tibble: 1,000 x 5
#>   a                   b              c     d e    
#>   <dttm>              <date>     <int> <dbl> <chr>
#> 1 2020-01-15 20:43:23 2020-01-22     1 0.368 n    
#> 2 2020-01-16 14:48:32 2020-01-27     2 0.612 l    
#> 3 2020-01-16 09:12:12 2020-02-06     3 0.415 p    
#> 4 2020-01-15 22:33:29 2020-02-05     4 0.212 m    
#> 5 2020-01-15 18:57:45 2020-02-02     5 0.733 i    
#> 6 2020-01-16 05:58:42 2020-01-29     6 0.460 n    
#> # … with 994 more rows


nycflights13::flights%>%
  print(n = 10, width = Inf)
```

```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

df$x # vector
df[['x']] # vector
df['x'] # 1 column df/tibble
df[[1]] # vector

# use placeholder . to use this sort of selection in pipes
df %>%
  .$x

df %>%
  .[['x']]

as.data.frame(df)%>%
  .['x']

as.data.frame(df)%>%
  .[['x']]
```

## How do you tell if an object is a tibble?
```{r}
mtcars

as_tibble(mtcars)
```

Can determine for sure with is_tibble or class(df), but can also tell by the formatting differences.
Tibble will display the variable type and just the head of the data.


## Compare the difference of the operations on a tibble vs a data.frame.
```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
class(df[, "xyz"])
class(df[, c("abc", "xyz")])

df2 <- as_tibble(df)
df2$x
class(df2[, "xyz"])
class(df2[, c("abc", "xyz")])
```

tibbles are more consistent with what type of object they return vs data.frames. This might mean you have to type more to get 
what you need, but saves bugs in the future.


## If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?
```{r}
var <- "mpg"

mtcars2 <- as_tibble(mtcars)

mtcars2[[var]]
```

Using $ will think you are looking for a variable named 'var' (df$var)


##Practice referring to non-syntactic names in the following data frame by:

Extracting the variable called 1.

Plotting a scatterplot of 1 vs 2.

Creating a new column called 3 which is 2 divided by 1.

Renaming the columns to one, two and three.
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)

annoying$`1`
annoying[[1]]
annoying[["1"]]


ggplot(annoying, aes(x = `1`, y = `2`))+
  geom_point()

annoying$`3` <- annoying$`2`/annoying$`1`

annoying %>%
  rename(one = `1`,
         two = `2`,
         three = `3`)
```


## What does tibble::enframe() do? When might you use it?
```{r}
??enframe

x <- runif(5)
y <- rnorm(5)

deframe(enframe(list(x = x, y = y)))

enframe(c(x = x, y = y))

enframe(x, name = "x")%>%
  bind_cols(enframe(y, name = "y"))
```
Converts  vectors or lists to data frames

## What option controls how many additional column names are printed at the footer of a tibble?

```{r}
mtcars2

?print.tbl()
```

n_extra controls additional column printing. Easy way to find this is to examine the print function for tibbles
