---
title: "vectors"
output: html_document
---

```{r}
library(tidyverse)
```

## Vector Basics
Two types of vectors:

1) Atomic: six types within this -- logical, integer, double, character, complex, raw. Integer and Double are 
also known as numeric vectors.
2) Lists: also called recursive vectors because lists can contain other lists.

Main difference between the two is atomic vectors are homogeneous while lists can be heterogeneous.

Vector has two key properties:
1) Type-- can be determined with typeof()
```{r}
typeof(letters)
typeof(1:10)
```

2) Length -- determined with length()
```{r}
length(letters)
```

Vectors can contain additional metadata in the form of *attributes*. These are used to build augmented vectors. 
Types of augmented vectors:

1) Factors -- built on top of integer vectors
2) Dates and date-times -- built on top of numeric vectors
3) Data Frames and Tibbles -- built on top of lists


## Atomic Vectors
Raw and complex are rarely used and not discussed below.

### Logical

Only take 3 values: TRUE, FALSE, NA. Usually created with comparison operators.

```{r}
1:10 %% 3 == 0
```


### Numeric

In R, numbers are doubles by default. Integers are made by placing an 'L' after the number.

```{r}
typeof(1)
typeof(1L)

```

Integer have NA as a special value.
Doubles have 4 special values: NA, NaN, -Inf, Inf.

Use helper functions to check for special values.
```{r}
is.finite()
is.infinite()
is.na()
is.nan()
```


### Character

### Missing Values
Each type of atomic vector has its own missing value. Normally you can just use NA and the conversion will
happen behind the scenes. Some functions are strict though so it's useful to know this is a thing.
```{r}
NA # Logical
NA_integer_ # integer
NA_real_ # double
NA_character_ # character
```

Exercises
1) Describe the difference between is.finite(x) and !is.infinite(x).
```{r}
x <- c(0, NA, NaN, -Inf, Inf)
is.finite(x)
!is.infinite(x)
```

2) Read the source code for dplyr::near() (Hint: to see the source code, drop the ()). How does it work?
```{r}
dplyr::near
```
Checks whether the numbers are within a tolerance. By default that tolerance is the sqrt of the smallest floating point number the computer
can represent.

4) Brainstorm at least four functions that allow you to convert a double to an integer. How do they differ? Be precise.
Can round to -Inf, Inf, towards zero, away from zero, to the nearest integer.

5) What functions from the readr package allow you to turn a string into logical, integer, and double vector?
```{r}
# parse functions allow this 

readr::parse_logical(c("TRUE", "FALSE", "1", "NA", "true"))
readr::parse_integer(c("123", "012", "NA"))
```
 

## Using Atomic Vectors

### Coercion

Explicit Coercion: this is calling something like as.double(), as.logical(), etc. Generally you should look 
earlier in your code to see if you could avoid the vector having the wrong type to begin with.

Implicit Coercion: happens when you use a vector in a specific context that expects a certain type of vector.
Eg. using a logical vector with a numeric summary function.

Implicit coercion is the trickier of the two--explicit coercion is easy to handle. 
```{r}
x <- sample(20, 100, replace = TRUE)

y <- x > 10

sum(y)

mean(y)
```

If you try to create a mixed atomic vector, the most complex type of item in the vector determines the type of the vector.
An atomic vector cannot have a mix of different types--the type is a property of the complete vector, not the components.

### Naming vectors
Can name any type of vector during creation with c() or after the fact with purrr::set_names(). This is most useful for 
subsetting.

```{r}
c(x = 1, y = 2, z = 3)

set_names(1:3, c("a", "b", "d"))
```

### Subsetting

Can subset a vector with [] using four types of things:

1) Numeric vector containing only integers.

```{r}
x <- c("one", "two", "three")
x[c(1, 3)]
x[c(1, 1, 1)]

x[c(-1)]
# cannot mix negative and positive integers
x[c(-1, 1)]

x[0]
```

2) Subsetting with a logical vector keeps all values corresponding to TRUE

```{r}
x <- c(1, 4, 10, 3, 7, NA, 8)

# all nonmissing values of x
x[!is.na(x)]

# all even values of x ( this will return NA values as well)

x[x %% 2 == 0]
```

3) Can subset named vector with a character vector

```{r}
x <- c(abc = 1, def = 2, xyz = 6)
x["abc"]
x[c("abc", "def")]
x[c("abc", "abc")]
```

4) Subsetting nothing ie x[], returns the complete x and is very useful for matrices.


[[ always extracts a single element(with names dropped) from the vector. Very useful for lists.

Exercises

1) What does mean(is.na(x)) tell you about a vector x? What about sum(!is.finite(x))?
mean(is.na(x)) tells you the proportion of the vector that is NA values. sum(!is.finite(x)) tells you the total
number of values in x that are infinite or NA.

```{r}
x <- c(1, 2, 3, 5, Inf, -Inf, 10, 11, NA, NA)

mean(is.na(x))
sum(!is.finite(x))
```


2) Carefully read the documentation of is.vector(). What does it actually test for? Why does is.atomic() not agree with the definition of atomic vectors above?
```{r}
?is.vector

is.vector(c(a = 1, b = 2, c= 3))

x <- 1:10
attr(x, "test") <- TRUE

is.vector(x)


```
Checking whether the object has any attributes others than names. If it doesn't, then it returns TRUE.
is.atomic() checks whether an object is one of the atomic types(eg logical). Will consider an object atomic
even if there are attributes attached. 

3) Compare and contrast setNames() with purrr::set_names().
```{r}
?setNames()
setNames(x)
?purrr::set_names()
purrr::set_names(x)
```
purrr::set_names has more features and stricter argument checking. purrr:set_names() will name itself if no
names are given. Can use functions to transform existing names with set_names(). Also has more explicit error
messaging--eg if your vector of names is shorter than the vector values you'll get an error with set_names(),
but setNames() will set missing names to NA silently.


4) Create functions that take a vector as input and returns:

The last value. Should you use [ or [[?
```{r}
x <- 1:10

last_value <- function(x){
  # should check for zero length vector
  if(length(x)){
    x[[length(x)]]
  }else{
    x
  }
}
last_value(c())
last_value(1:100)
```

The elements at even numbered positions.
```{r}
even_values <- function(x){
  if(length(x)){
    x[seq_along(x) %% 2 == 0]
  }else{
    x
  }
}

even_values(1:10)
even_values(1:11)
even_values(numeric())
even_values(1)
even_values(letters)
```


Every element except the last value.
```{r}
remove_last <- function(x){
  if(length(x)){
    x[-length(x)]
  }else{
    x
  }
}

remove_last(1:10)
remove_last(c("a", "b", "c"))
remove_last(c(TRUE, FALSE, FALSE))

remove_last(numeric())
remove_last(1)
```


Only even numbers (and no missing values).
```{r}
even_numbers <- function(x){
  x[(x %% 2 == 0) & !is.nan(x) & !is.infinite(x)]
}

even_numbers(c(1, 2, 3, 4, 5, NA, NaN, Inf))
```


5) Why is x[-which(x > 0)] not the same as x[x <= 0]?
```{r}
x <- c(-5:5, NA, NaN, Inf, -Inf)

x[-which(x>0)]
x[x<=0]
```


6) What happens when you subset with a positive integer that’s bigger than the length of the vector? What happens when you subset with a name that doesn’t exist?

```{r}
x <- 1:10

x[111]

x <- c(a = 10, b = 11)
x["c"]
```

## Recursive Vectors(LISTS)

```{r}
x <- list(1, 2, 3)

x
str(x)

x <- list(a = 1, b = 2, c = 3)
str(x)

y <- list("a", 1L, 1.5, TRUE)
str(y)

z <- list(list(1, 2), list(3, 4))
str(z)

```

### Subsetting Lists

```{r}
a <- list(a = 1:3, b = "a string", c = pi, d = list(-1, -5))
str(a)


```

[ extracts a sub-list. The result will always be a list. Just like with vectors you can subset with logical, integer
or character vector

```{r}
typeof(a[1])
str(a[1])
```

[[ extracts a single component from the list. It removes a level of hierarchy from the list.

```{r}
str(a[1])
str(a[[1]])
```

$ is shorthand for extracting named elements of a list. It works similarly to [[ except that you don't need to use quotes.

```{r}
str(a$a)
str(a[["a"]])

```

[ vs [[ is important for lists. [[ drills into a list while [ returns a new, smaller list.


Exercises

1) Draw the following lists as nested sets:
    list(a, b, list(c, d), list(e, f))
    list(list(list(list(list(list(a))))))
    
```{r}

```
    

What happens if you subset a tibble as if you’re subsetting a list? What are the key differences between a list and a tibble?

```{r}
tib <- tibble(
  x = 1:4,
  y = c("a", "b", "c", "d")
)

tib$x
tib["x"]
tib[["x"]][2]
```

Main difference of a tibble is that each element(column) of the tibble must be the same length(number of rows). Lists can have 
vectors with different lengths.

## Attributes

Any vector can contain additional metadata through attributes(named list of vectors that can be attached to any object).

Three important attributes:
1) Names: used to name the elements of a vector
2) Dimensions: make a vector behave like a matrix or array.
3) Class: used to implement the S3 object oriented system.

## Augmented Vectors

Vectors with additional attributes--including class. Things like factors, dates, date-times, and tibbles would go here.

### Factors
Built ontop of integers. They are designed to represent categorical data that can take a finite set of possible values.
Have a levels attribute.

### Dates and Date-times

Dates are a numeric vector that represents the number of days since 1 Jan. 1970.

Date-times are numeric vectors with class POSIXct that represent the number of seconds since 1 Jan. 1970.

### Tibbles

Augmented lists. Have tbl_df, tbl, data.frame, names and row.names attributes.

```{r}
tb <- tibble::tibble(x = 1:5, y = 5:1)

typeof(tb)
attributes(tb)
```


Exercises

1) What does hms::hms(3600) return? How does it print? What primitive type is the augmented vector built on top of? What attributes does it use?
```{r}
typeof(hms::hms(3600))
attributes(hms::hms(3600))
```

Returns 1 hour. Built on top of a double with attributes units, hms, and difftime.

2) Try and make a tibble that has columns with different lengths. What happens?
```{r}
tibble::tibble(x = 1:4, y = 1:2)
```


3) Based on the definition above, is it ok to have a list as a column of a tibble?

```{r}
tibble(x = list(1, 2, 3, "a"), 
       y = 1:4)
```

