---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Helper packages
library(dplyr)    # for general data wrangling needs

# Modeling packages
library(gbm)      # for original implementation of regular and stochastic GBMs
library(h2o)      # for a java-based implementation of GBM variants
library(xgboost)  # for fitting extreme gradient boosting
library(rsample)

set.seed(123)
ames <- AmesHousing::make_ames()

#splitting data into stratified random sample with rsample
split <- initial_split(ames, prop = .7,
                       strata = 'Sale_Price')
# create the train and test dfs
ames_train <- training(split)
ames_test <- testing(split)

# initialize all h2o stuff
h2o.init(max_mem_size = '10g')

train_h2o <- as.h2o(ames_train)
response <- "Sale_Price"
predictors <- setdiff(colnames(ames_train), response)
```

Can use formula interface (gbm::gbm) or x y specification interface(gbm::gbm.fit), but gbm.fit is more 
efficient
```{r}

# Step 1 and 2: set high learning rate and then find optimal number of trees
# run basic gbm
set.seed(123)
ames_gbm1 <- gbm(
  formula = Sale_Price ~ .,
  data = ames_train,
  distribution = "gaussian", # SSE loss function
  n.trees = 5000,
  shrinkage = .1,
  interaction.depth = 3, 
  n.minobsinnode = 10,
  cv.folds = 10
)

# find index for number of trees with minimum CV error
best <- which.min(ames_gbm1$cv.error)

# get MSE and compute RMSE
sqrt(ames_gbm1$cv.error[best])

# plot error curve
gbm.perf(ames_gbm1, method = 'cv')
```

## Tuning
These various models take a while to train. Don't run unless you are going to use them.
```{r}

# Step 3: Fix tree parameters and tune learning rate
# create grid search
hyper_grid <- expand.grid(
  learning_rate = c(0.3, 0.1, 0.05, 0.01, 0.005),
  RMSE = NA,
  trees = NA,
  time = NA
)

# execute grid search
for(i in seq_len(nrow(hyper_grid))) {

  # fit gbm
  set.seed(123)  # for reproducibility
  train_time <- system.time({
    m <- gbm(
      formula = Sale_Price ~ .,
      data = ames_train,
      distribution = "gaussian",
      n.trees = 5000, 
      shrinkage = hyper_grid$learning_rate[i], 
      interaction.depth = 3, 
      n.minobsinnode = 10,
      cv.folds = 10 
   )
  })
  
  # add SSE, trees, and training time to results
  hyper_grid$RMSE[i]  <- sqrt(min(m$cv.error))
  hyper_grid$trees[i] <- which.min(m$cv.error)
  hyper_grid$Time[i]  <- train_time[["elapsed"]]

}

arrange(hyper_grid, RMSE)
```


```{r}
# Step 4: fix learning rate and tune tree specific parameters for that rate
# search grid
hyper_grid <- expand.grid(
  n.trees = 6000,
  shrinkage = 0.01,
  interaction.depth = c(3, 5, 7),
  n.minobsinnode = c(5, 10, 15)
)

# create model fit function
model_fit <- function(n.trees, shrinkage, interaction.depth, n.minobsinnode) {
  set.seed(123)
  m <- gbm(
    formula = Sale_Price ~ .,
    data = ames_train,
    distribution = "gaussian",
    n.trees = n.trees,
    shrinkage = shrinkage,
    interaction.depth = interaction.depth,
    n.minobsinnode = n.minobsinnode,
    cv.folds = 10
  )
  # compute RMSE
  sqrt(min(m$cv.error))
}

# perform search grid with functional programming
hyper_grid$rmse <- purrr::pmap_dbl(
  hyper_grid,
  ~ model_fit(
    n.trees = ..1,
    shrinkage = ..2,
    interaction.depth = ..3,
    n.minobsinnode = ..4
    )
)
# results
arrange(hyper_grid, rmse)
```

## Stochastic GBM Implementation
```{r}
# refined hyperparameter grid
hyper_grid <- list(
  sample_rate = c(0.5, 0.75, 1),              # row subsampling
  col_sample_rate = c(0.5, 0.75, 1),          # col subsampling for each split
  col_sample_rate_per_tree = c(0.5, 0.75, 1)  # col subsampling for each tree
)

# random grid search strategy
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "mse",
  stopping_tolerance = 0.001,   
  stopping_rounds = 10,         
  max_runtime_secs = 60*60      
)

# perform grid search 
grid <- h2o.grid(
  algorithm = "gbm",
  grid_id = "gbm_grid",
  x = predictors, 
  y = response,
  training_frame = train_h2o,
  hyper_params = hyper_grid,
  ntrees = 6000,
  learn_rate = 0.01,
  max_depth = 7,
  min_rows = 5,
  nfolds = 10,
  stopping_rounds = 10,
  stopping_tolerance = 0,
  search_criteria = search_criteria,
  seed = 123
)

# collect the results and sort by our model performance metric of choice
grid_perf <- h2o.getGrid(
  grid_id = "gbm_grid", 
  sort_by = "mse", 
  decreasing = FALSE
)

grid_perf

# Grab the model_id for the top model, chosen by cross validation error
best_model_id <- grid_perf@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)

# Now let’s get performance metrics on the best model
h2o.performance(model = best_model, xval = TRUE)
```

## XGBOOST

```{r}
library(recipes)
xgb_prep <- recipe(Sale_Price ~ ., data = ames_train) %>%
  step_integer(all_nominal()) %>%
  prep(training = ames_train, retain = TRUE) %>%
  juice()

X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "Sale_Price")])
Y <- xgb_prep$Sale_Price

set.seed(123)
ames_xgb <- xgb.cv(
  data = X,
  label = Y,
  nrounds = 6000,
  objective = "reg:linear",
  early_stopping_rounds = 50, 
  nfold = 10,
  params = list(
    eta = 0.1,
    max_depth = 3,
    min_child_weight = 3,
    subsample = 0.8,
    colsample_bytree = 1.0),
  verbose = 0
)  

# minimum test CV RMSE
min(ames_xgb$evaluation_log$test_rmse_mean)
```

## Tuning strategy
```{r}
library(recipes)
# encode categorical variables, prep the recipe and then juice--actually do--the preprocessing-- in this case encoding
xgb_prep <- recipe(Sale_Price ~ . , data = ames_train)%>%
  step_integer(all_nominal())%>%
  prep(training = ames_train, retain = TRUE)%>%
  juice()

# grab features from tibble
X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "Sale_Price")])
# grab response from tibble
Y <- xgb_prep$Sale_Price

# grid search as done in other sections was used to find hyperparameters used below
set.seed(123)

ames_xgb <- xgb.cv(
  data = X,
  label = Y,
  nrounds = 6000,
  objective = 'reg:linear',
  early_stopping_rounds = 50,
  nfold = 10,
  params = list(
    eta = 0.1,
    max_depth = 3,
    min_child_weight = 3,
    subsample = 0.8,
    colsample_bytree = 1.0),
  verbose = 0
)

# min test CV RMSE
min(ames_xgb$evaluation_log$test_rmse_mean)
```


Now you can assess if the model is overfitting which could be impacting performance. Do this with a grid search
that explores the regularization parameters space.
```{r}
# hyperparameter grid. see xgb.train for explaination of each parameter
hyper_grid <- expand.grid(
  eta = 0.01,
  max_depth = 3, 
  min_child_weight = 3,
  subsample = 0.5, 
  colsample_bytree = 0.5,
  gamma = c(0, 1, 10, 100, 1000),
  lambda = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  alpha = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  rmse = 0,          # a place to dump RMSE results
  trees = 0          # a place to dump required number of trees
)

# do the grid search
for(i in seq_len(nrow(hyper_grid))){
  set.seed(123)
  m <- xgb.cv(
    data = X,
    label = Y,
    nrounds = 4000,
    objective = 'reg:linear',
    early_stopping_rounds = 50,
    nfold = 10,
    verbose = 0,
    params = list(
      eta = hyper_grid$eta[i],
      max_depth = hyper_grid$max_depth[i],
      min_child_weight = hyper_grid$min_child_weight[i],
      subsample = hyper_grid$subsample[i],
      colsample_bytree = hyper_grid$colsample_bytree[i],
      gamma = hyper_grid$gamma[i],
      lambda = hyper_grid$lambda[i],
      alpha = hyper_grid$alpha[i]
    )
  )
  hyper_grid$rmse[i] <- min(m$evaluation_log$test_rmse_mean)
  hyper_grid$trees[i] <- m$best_iteration
  
}

# results
hyper_grid %>%
  filter(rmse > 0)%>%
  arrange(rmse)%>%
  glimpse()
```

Once you have the optimal hyperparameters, fit the final model with xgb.train (or xgboost)
```{r}
# optimal parameter list
params <- list(
  eta = 0.01,
  max_depth = 3,
  min_child_weight = 3,
  subsample = 0.5,
  colsample_bytree = 0.5
)

# train final model
xgb.fit.final <- xgboost(
  params = params,
  data = X,
  label = Y,
  nrounds = 3944,
  objective = "reg:linear",
  verbose = 0
)
```

## Feature Importance
Can use VIP for feature importance. Default method is gain, but that can be changed with 'type' argument.
```{r}
# variable importance plot
vip::vip(xgb.fit.final)

# can also use xgboost.
xgb.ggplot.importance(xgb.importance(xgb.fit.final))


```

